---
title: 'CIND 820 Initial Results and the Code'
output:
  html_document: default
  pdf_document: default
  word_document: default
---
<center> <h1> Initial Results and the Code (10%) </h1> </center>
<center>  <h3> Ryan Bursey </h2> </center>
<center> <h3> DA0 & 500732132 </h2> </center>
---

#### 0. Install Packages
```{r}
##text to knit document, remove ## to install

##install.packages("corrplot")
##install.packages("ROSE")
##install.packages("randomForest")
##install.packages("caret")
##install.packages("ggplot2")
##install.packages("rpart.plot")
```

#### 1. Reading the bank dataset
```{r}
bank_data <- read.csv("C:\\Users\\rburs\\Documents\\Bank Churn - Raw File.csv",
                       header = T)
```

#### 2. Check the data types of the attributes
```{r}
str(bank_data)
```


#### 3. Remove Irrelevant Columms

```{r}

bank_data <- bank_data[4:14]

```

#### 4.  Encode Categorical Data

```{r}
bank_data$Gender <- as.factor(bank_data$Gender)

levels(bank_data$Gender)

bank_data$Geography <- as.factor(bank_data$Geography)

levels(bank_data$Geography)

str(bank_data)

```

#### 5. Check the dataset for any missing values
```{r}
which(is.na(bank_data[1]))
which(is.na(bank_data[2]))
which(is.na(bank_data[3]))
which(is.na(bank_data[4]))
which(is.na(bank_data[5]))
which(is.na(bank_data[6]))
which(is.na(bank_data[7]))
which(is.na(bank_data[8]))
which(is.na(bank_data[9]))
which(is.na(bank_data[10]))
which(is.na(bank_data[11]))

print("No missing values in this dataset")

```
#### 6. Summary Measures


```{r}
summary(bank_data)

```

#### 7. Check the Variable Correlation?
```{r}
library(corrplot)

corr_data <- cbind.data.frame(bank_data[1],bank_data[4:11])

colnames(corr_data) <- c("CreditScore",
                         "Age",
                         "Tenure",
                         "Balance",
                         "NumOfProducts",
                         "HasCrCard",
                         "IsActiveMember",
                         "EstimatedSalary",
                         "Exited")

correlation_calc <- cor(corr_data)

corrplot(correlation_calc, type="upper",)

correlation_calc

print("Based on this, there is little correlation between most of the columns. Main items of note are that Number of Products and Balance have a weak to moderate negative correlation and that there is a weak to moderate positive correlation between Age and Exited. Is active member and exited also appears to have a weak negative correlation.")

```

#### 8. Graphing the frequency distribution of the variables
```{r}
par(mfrow=c(3,2))

##Distribution of Credit Score
hist(bank_data$CreditScore, 
     xlab = "Credit Score", 
     main = "Bank Customer Distributions")

##Distribution of Age
hist(bank_data$Age, 
     xlab = "Age", 
     main = "Bank Customer Distributions")

##Distribution of Tenure
hist(bank_data$Tenure, 
     xlab = "Tenure",
     main = "Bank Customer Distributions")

##Distribution of Balance
hist(bank_data$Balance, 
     xlab = "Balance", 
     main = "Bank Customer Distributions")

##Distribution of Number of Products
hist(bank_data$NumOfProducts, 
     xlab = "Number of Products", 
     main = "Bank Customer Distributions")

##Distribution of Estimated Salary
hist(bank_data$EstimatedSalary, 
     xlab = "Estimated Salary", 
     main = "Bank Customer Distributions")

print("Credit score appears to have a relatively normal distribution. Age is right skewed with most customers aged 30-40. Tenure has an equal distribution. Balance contains a large portion of customers with a 0 balance but is centered around 125K. Most customers have either 1 or 2 produts. Estimated salary has an equal distribution.")

```

#### 9. Count Frequency of other data types
```{r}

barplot(prop.table(table(bank_data$Exited)),
        col = rainbow(2),
        main = "Class Distribution",
        ylim = c(0,1))

barplot(prop.table(table(bank_data$Geography)),
        col = rainbow(3),
        main = "Geography Distribution",
        ylim = c(0,1))

barplot(prop.table(table(bank_data$Gender)),
        col = rainbow(3),
        main = "Gender Distribution",
        ylim = c(0,1))

barplot(prop.table(table(bank_data$HasCrCard)),
        col = rainbow(3),
        main = "Has Credit Card Distribution",
        ylim = c(0,1))


barplot(prop.table(table(bank_data$IsActiveMember)),
        col = rainbow(3),
        main = "IsActiveMember Distribution",
        ylim = c(0,1))
print("We have a class imbalance, as about 80% of our dataset has an exited value of 0. About 50% of customers are from France. This data set has slightly more males than females. About 75% of the data set has a credit card. Active members are about as frequent as non-active members.")

```


#### 10. Boxplots of Variables

```{r}

##Boxplot of Credit Score
boxplot(bank_data$CreditScore, 
     xlab = "Credit Score", 
     main = "Bank Customer Boxplots",
     horizontal = TRUE)

##Boxplot of Age
boxplot(bank_data$Age, 
        xlab = "Age", 
        main = "Bank Customer Boxplots",
        horizontal = TRUE)

##Boxplot of Tenure
boxplot(bank_data$Tenure, 
        xlab = "Tenure", 
        main = "Bank Customer Boxplots",
        horizontal = TRUE)

##Boxplot of Balance
boxplot(bank_data$Balance, 
        xlab = "Balance", 
        main = "Bank Customer Boxplots",
        horizontal = TRUE)

##Boxplot of Number of Products
boxplot(bank_data$NumOfProducts, 
        xlab = "Number of Products", 
        main = "Bank Customer Boxplots", 
        horizontal = TRUE)

##Boxplot of Estimated Salary
boxplot(bank_data$EstimatedSalary, 
        xlab = "Estimated Salary", 
        main = "Bank Customer Boxplots",
        horizontal = TRUE)

print("Looking at the Boxplots, we can infer that there are potential outliers in Credit Score, Age and Number of Products")
```

#### 11. Class balancing vis ROSE Oversampling
```{r}
library(ROSE)

over <- ovun.sample(Exited~., data = bank_data, method = "over",seed = 1, N = (10000+(7963-2037)))$data
table(over$Exited)

```

#### 12. Class balancing vis ROSE Undersampling
```{r}
library(ROSE)

under <- ovun.sample(Exited~., data = bank_data, method = "under",seed = 1, N = (10000-(7963-2037)))$data
table(under$Exited)
```

####  13. Class balancing vis ROSE Bothsampling
```{r}
library(ROSE)

both <- ovun.sample(Exited~., data = bank_data, method = "both",seed = 1, N = 10000)$data
table(both$Exited)
```

#### 14. Remove Outliers
```{r}

outliers <- function(x) {

  Q1 <- quantile(x, probs=.25)
  Q3 <- quantile(x, probs=.75)
  iqr = Q3-Q1

 upper_limit = Q3 + (iqr*1.5)
 lower_limit = Q1 - (iqr*1.5)

 x > upper_limit | x < lower_limit
}

remove_outliers <- function(df, cols = names(df)) {
  for (col in cols) {
    df <- df[!outliers(df[[col]]),]
  }
  df
}

outliers_bank <- remove_outliers(bank_data,c("CreditScore","Age","Tenure","Balance","NumOfProducts","EstimatedSalary"))
outliers_over <- remove_outliers(over,c("CreditScore","Age","Tenure","Balance","NumOfProducts","EstimatedSalary"))
outliers_under <- remove_outliers(under,c("CreditScore","Age","Tenure","Balance","NumOfProducts","EstimatedSalary"))
outliers_both <- remove_outliers(both,c("CreditScore","Age","Tenure","Balance","NumOfProducts","EstimatedSalary"))

```

#### 15. Normalize Numeric Attributes
```{r}
min_max_norm <- function(x) {(x - min(x)) / (max(x) - min(x))}

##Normalize bank_data data set
bank_norm_data1 <- as.data.frame(lapply(bank_data[1],min_max_norm))

bank_norm_data2 <- as.data.frame(lapply(bank_data[4:10],min_max_norm))

bank_norm <- cbind.data.frame(bank_norm_data1,bank_norm_data2,Geography = bank_data$Geography, Gender = bank_data$Gender, Exited = bank_data$Exited)

##Normalize over data set
over_norm_data1 <- as.data.frame(lapply(over[1],min_max_norm))

over_norm_data2 <- as.data.frame(lapply(over[4:10],min_max_norm))

over_norm <- cbind.data.frame(over_norm_data1,over_norm_data2,Geography = over$Geography, Gender = over$Gender, Exited = over$Exited)

##Normalize under data set
under_norm_data1 <- as.data.frame(lapply(under[1],min_max_norm))

under_norm_data2 <- as.data.frame(lapply(under[4:10],min_max_norm))

under_norm <- cbind.data.frame(under_norm_data1,under_norm_data2,Geography = under$Geography, Gender = under$Gender, Exited = under$Exited)

##Normalize both data set
both_norm_data1 <- as.data.frame(lapply(both[1],min_max_norm))

both_norm_data2 <- as.data.frame(lapply(both[4:10],min_max_norm))

both_norm <- cbind.data.frame(both_norm_data1,both_norm_data2,Geography = both$Geography, Gender = both$Gender, Exited = both$Exited)

##Normalize outliers_bank data set
outliers_norm_data1 <- as.data.frame(lapply(outliers_bank[1],min_max_norm))

outliers_norm_data2 <- as.data.frame(lapply(outliers_bank[4:10],min_max_norm))

outliers_norm <- cbind.data.frame(outliers_norm_data1,outliers_norm_data2,Geography = outliers_bank$Geography, Gender = outliers_bank$Gender, Exited = outliers_bank$Exited)

##Normalize outliers_over data set
outliers_over_norm_data1 <- as.data.frame(lapply(outliers_over[1],min_max_norm))

outliers_over_norm_data2 <- as.data.frame(lapply(outliers_over[4:10],min_max_norm))

outliers_over_norm <- cbind.data.frame(outliers_over_norm_data1,outliers_over_norm_data2,Geography = outliers_over$Geography, Gender = outliers_over$Gender, Exited = outliers_over$Exited)

##Normalize outliers_under data set
outliers_under_norm_data1 <- as.data.frame(lapply(outliers_under[1],min_max_norm))

outliers_under_norm_data2 <- as.data.frame(lapply(outliers_under[4:10],min_max_norm))

outliers_under_norm <- cbind.data.frame(outliers_under_norm_data1,outliers_under_norm_data2,Geography = outliers_under$Geography, Gender = outliers_under$Gender, Exited = outliers_under$Exited)

##Normalize outliers_both data set
outliers_both_norm_data1 <- as.data.frame(lapply(outliers_both[1],min_max_norm))

outliers_both_norm_data2 <- as.data.frame(lapply(outliers_both[4:10],min_max_norm))

outliers_both_norm <- cbind.data.frame(outliers_both_norm_data1,outliers_both_norm_data2,Geography = outliers_both$Geography, Gender = outliers_both$Gender, Exited = outliers_both$Exited)

```

#### 16. Divide the dataset to training and test sets
```{r}

##data partition for bank_data
set.seed(222)
ind <- sample(2, nrow(bank_data), replace = TRUE, prob = c(0.7, 0.3))
bank_data_train <- bank_data[ind==1,]
bank_data_test <- bank_data[ind==2,]

##data partition for bank_norm
set.seed(222)
ind <- sample(2, nrow(bank_norm), replace = TRUE, prob = c(0.7, 0.3))
bank_norm_train <- bank_norm[ind==1,]
bank_norm_test <- bank_norm[ind==2,]

##data partition for outliers_bank
set.seed(222)
ind <- sample(2, nrow(outliers_bank), replace = TRUE, prob = c(0.7, 0.3))
outliers_bank_train <- outliers_bank[ind==1,]
outliers_bank_test <- outliers_bank[ind==2,]

##data partition for outliers_norm
set.seed(222)
ind <- sample(2, nrow(outliers_norm), replace = TRUE, prob = c(0.7, 0.3))
outliers_norm_train <- outliers_norm[ind==1,]
outliers_norm_test <- outliers_norm[ind==2,]

##data partition for over
set.seed(222)
ind <- sample(2, nrow(over), replace = TRUE, prob = c(0.7, 0.3))
over_train <- over[ind==1,]
over_test <- over[ind==2,]

##data partition for over_norm
set.seed(222)
ind <- sample(2, nrow(over_norm), replace = TRUE, prob = c(0.7, 0.3))
over_norm_train <- over_norm[ind==1,]
over_norm_test <- over_norm[ind==2,]

##data partition for outliers_over
set.seed(222)
ind <- sample(2, nrow(outliers_over), replace = TRUE, prob = c(0.7, 0.3))
outliers_over_train <- outliers_over[ind==1,]
outliers_over_test <- outliers_over[ind==2,]

##data partition for outliers_over_norm
set.seed(222)
ind <- sample(2, nrow(outliers_over_norm), replace = TRUE, prob = c(0.7, 0.3))
outliers_over_norm_train <- outliers_over_norm[ind==1,]
outliers_over_norm_test <- outliers_over_norm[ind==2,]

##data partition for under
set.seed(222)
ind <- sample(2, nrow(under), replace = TRUE, prob = c(0.7, 0.3))
under_train <- under[ind==1,]
under_test <- under[ind==2,]

##data partition for under_norm
set.seed(222)
ind <- sample(2, nrow(under_norm), replace = TRUE, prob = c(0.7, 0.3))
under_norm_train <- under_norm[ind==1,]
under_norm_test <- under_norm[ind==2,]

##data partition for outliers_under
set.seed(222)
ind <- sample(2, nrow(outliers_under), replace = TRUE, prob = c(0.7, 0.3))
outliers_under_train <- outliers_under[ind==1,]
outliers_under_test <- outliers_under[ind==2,]

##data partition for outliers_under_norm
set.seed(222)
ind <- sample(2, nrow(outliers_under_norm), replace = TRUE, prob = c(0.7, 0.3))
outliers_under_norm_train <- outliers_under_norm[ind==1,]
outliers_under_norm_test <- outliers_under_norm[ind==2,]

##data partition for both
set.seed(222)
ind <- sample(2, nrow(both), replace = TRUE, prob = c(0.7, 0.3))
both_train <- both[ind==1,]
both_test <- both[ind==2,]

##data partition for both_norm
set.seed(222)
ind <- sample(2, nrow(both_norm), replace = TRUE, prob = c(0.7, 0.3))
both_norm_train <- both_norm[ind==1,]
both_norm_test <- both_norm[ind==2,]

##data partition for outliers_both
set.seed(222)
ind <- sample(2, nrow(outliers_both), replace = TRUE, prob = c(0.7, 0.3))
outliers_both_train <- outliers_both[ind==1,]
outliers_both_test <- outliers_both[ind==2,]

##data partition for outliers_both_norm
set.seed(222)
ind <- sample(2, nrow(outliers_both_norm), replace = TRUE, prob = c(0.7, 0.3))
outliers_both_norm_train <- outliers_both_norm[ind==1,]
outliers_both_norm_test <- outliers_both_norm[ind==2,]

```

#### 17A. Model Bank_Data
```{r}
library(rpart)
library(rpart.plot)
fit <- rpart(Exited~., data = bank_data_train , method = 'class')

rpart.plot(fit, extra = 106)

predict_exited <-predict(fit, bank_data_test, type = 'class')

table_mat <- table(bank_data_test$Exited, predict_exited)
table_mat

accuracy_Test <- sum(diag(table_mat)) / sum(table_mat)
print(paste('Accuracy for test', accuracy_Test))
```

#### 17B. Model Bank_Norm
```{r}
library(rpart)
library(rpart.plot)
fit <- rpart(Exited~., data = bank_norm_train , method = 'class')

rpart.plot(fit, extra = 106)

predict_exited <-predict(fit, bank_norm_test, type = 'class')

table_mat <- table(bank_data_test$Exited, predict_exited)
table_mat

accuracy_Test <- sum(diag(table_mat)) / sum(table_mat)
print(paste('Accuracy for test', accuracy_Test))
```

#### 17C. Model outliers_bank
```{r}
library(rpart)
library(rpart.plot)
fit <- rpart(Exited~., data = outliers_bank_train , method = 'class')

rpart.plot(fit, extra = 106)

predict_exited <-predict(fit, outliers_bank_test, type = 'class')

table_mat <- table(outliers_bank_test$Exited, predict_exited)
table_mat

accuracy_Test <- sum(diag(table_mat)) / sum(table_mat)
print(paste('Accuracy for test', accuracy_Test))

 
```

#### 17D. Model outliers_norm
```{r}
library(rpart)
library(rpart.plot)
fit <- rpart(Exited~., data = outliers_norm_train , method = 'class')

rpart.plot(fit, extra = 106)

predict_exited <-predict(fit, outliers_norm_test, type = 'class')

table_mat <- table(outliers_norm_test$Exited, predict_exited)
table_mat

accuracy_Test <- sum(diag(table_mat)) / sum(table_mat)
print(paste('Accuracy for test', accuracy_Test))

 
```

#### 17E. Model over
```{r}
library(rpart)
library(rpart.plot)
fit <- rpart(Exited~., data = over_train , method = 'class')

rpart.plot(fit, extra = 106)

predict_exited <-predict(fit, over_test, type = 'class')

table_mat <- table(over_test$Exited, predict_exited)
table_mat

accuracy_Test <- sum(diag(table_mat)) / sum(table_mat)
print(paste('Accuracy for test', accuracy_Test))

 
```

#### 17F. Model over_norm
```{r}
library(rpart)
library(rpart.plot)
fit <- rpart(Exited~., data = over_norm_train , method = 'class')

rpart.plot(fit, extra = 106)

predict_exited <-predict(fit, over_norm_test, type = 'class')

table_mat <- table(over_norm_test$Exited, predict_exited)
table_mat

accuracy_Test <- sum(diag(table_mat)) / sum(table_mat)
print(paste('Accuracy for test', accuracy_Test))

 
```

#### 17G. Model outliers_over
```{r}
library(rpart)
library(rpart.plot)
fit <- rpart(Exited~., data = outliers_over_train , method = 'class')

rpart.plot(fit, extra = 106)

predict_exited <-predict(fit, outliers_over_test, type = 'class')

table_mat <- table(outliers_over_test$Exited, predict_exited)
table_mat

accuracy_Test <- sum(diag(table_mat)) / sum(table_mat)
print(paste('Accuracy for test', accuracy_Test))

 
```

#### 17H. Model outliers_over_norm
```{r}
library(rpart)
library(rpart.plot)
fit <- rpart(Exited~., data = outliers_over_norm_train , method = 'class')

rpart.plot(fit, extra = 106)

predict_exited <-predict(fit, outliers_over_norm_test, type = 'class')

table_mat <- table(outliers_over_norm_test$Exited, predict_exited)
table_mat

accuracy_Test <- sum(diag(table_mat)) / sum(table_mat)
print(paste('Accuracy for test', accuracy_Test))

 
```

#### 17I. Model under
```{r}
library(rpart)
library(rpart.plot)
fit <- rpart(Exited~., data = under_train , method = 'class')

rpart.plot(fit, extra = 106)

predict_exited <-predict(fit, under_test, type = 'class')

table_mat <- table(under_test$Exited, predict_exited)
table_mat

accuracy_Test <- sum(diag(table_mat)) / sum(table_mat)
print(paste('Accuracy for test', accuracy_Test))
 
```

#### 17J. Model under_norm
```{r}
library(rpart)
library(rpart.plot)
fit <- rpart(Exited~., data = under_norm_train , method = 'class')

rpart.plot(fit, extra = 106)

predict_exited <-predict(fit, under_norm_test, type = 'class')

table_mat <- table(under_norm_test$Exited, predict_exited)
table_mat

accuracy_Test <- sum(diag(table_mat)) / sum(table_mat)
print(paste('Accuracy for test', accuracy_Test))
 
```

#### 17K. Model outliers_under
```{r}
library(rpart)
library(rpart.plot)
fit <- rpart(Exited~., data = outliers_under_train , method = 'class')

rpart.plot(fit, extra = 106)

predict_exited <-predict(fit, outliers_under_test, type = 'class')

table_mat <- table(outliers_under_test$Exited, predict_exited)
table_mat

accuracy_Test <- sum(diag(table_mat)) / sum(table_mat)
print(paste('Accuracy for test', accuracy_Test))
 
```

#### 17L. Model outliers_under_norm
```{r}
library(rpart)
library(rpart.plot)
fit <- rpart(Exited~., data = outliers_under_norm_train , method = 'class')

rpart.plot(fit, extra = 106)

predict_exited <-predict(fit, outliers_under_norm_test, type = 'class')

table_mat <- table(outliers_under_norm_test$Exited, predict_exited)
table_mat

accuracy_Test <- sum(diag(table_mat)) / sum(table_mat)
print(paste('Accuracy for test', accuracy_Test))
 
```

#### 17M. Model both
```{r}
library(rpart)
library(rpart.plot)
fit <- rpart(Exited~., data = both_train , method = 'class')

rpart.plot(fit, extra = 106)

predict_exited <-predict(fit, both_test, type = 'class')

table_mat <- table(both_test$Exited, predict_exited)
table_mat

accuracy_Test <- sum(diag(table_mat)) / sum(table_mat)
print(paste('Accuracy for test', accuracy_Test))
 
```

#### 17N. Model both_norm
```{r}
library(rpart)
library(rpart.plot)
fit <- rpart(Exited~., data = both_norm_train , method = 'class')

rpart.plot(fit, extra = 106)

predict_exited <-predict(fit, both_norm_test, type = 'class')

table_mat <- table(both_norm_test$Exited, predict_exited)
table_mat

accuracy_Test <- sum(diag(table_mat)) / sum(table_mat)
print(paste('Accuracy for test', accuracy_Test))
 
```

#### 17N. Model outliers_both
```{r}
library(rpart)
library(rpart.plot)
fit <- rpart(Exited~., data = outliers_both_train , method = 'class')

rpart.plot(fit, extra = 106)

predict_exited <-predict(fit, outliers_both_test, type = 'class')

table_mat <- table(outliers_both_test$Exited, predict_exited)
table_mat

accuracy_Test <- sum(diag(table_mat)) / sum(table_mat)
print(paste('Accuracy for test', accuracy_Test))
 
```

#### 17N. Model outliers_both_norm
```{r}
library(rpart)
library(rpart.plot)
fit <- rpart(Exited~., data = outliers_both_norm_train , method = 'class')

rpart.plot(fit, extra = 106)

predict_exited <-predict(fit, outliers_both_norm_test, type = 'class')

table_mat <- table(outliers_both_norm_test$Exited, predict_exited)
table_mat

accuracy_Test <- sum(diag(table_mat)) / sum(table_mat)
print(paste('Accuracy for test', accuracy_Test))
 
```

#### 18. Final Initial Code Comments
```{r}

print('We get the highest accuracy using the base bank dataset, but knowing that there is a class imbalance we cannot make any confident conclusions of the models accuracy without class balancing. Of all models that incorporated some type of class balancing strategy, the best decision tree does not remove outliers and utilizes a combination of under and over sampling methods to balance the classes. The dataset: both_test gives us an accuracy of 75.7%. In future iterations we will compare the results of this model vs. other methods.')
 
```
